Tabular Constraint Learning

### 2016-04-12 Paper/ECAI/tabular.tex

\newcommand{\ecrank}[2]{\ensuremath{#1 = \mathit{RANK}(#2)}}

@@

\section{Approach to Tabular Constraint Learning}
As we have seen the statement before assumes the set of groups to be given. It is often, however, not the case and a possible approach is to extract the groups from the CSV file directly. In our system, we have implement the group construction step using the type-consistency check. \sergey{Samuel, we need details here, I think} In general, separating data from a header is a task on its own  \cite{header} and goes beyond the scope of this paper.

@@

Now we describe how the problem of Condensed Tabular Constraint Learning can be solved. The algorithm mimics the structure of the problem in the following way: first we need to order constraints in the order of specificity, then for each constraint we need to generate possible candidate groups, and in the next step we find the subgroups satisfying the constrain. After an iteration we accumulate all solutions and proceed to the next constraint.

@@

\samuel{Running example (RANK?)}
\paragraph{Constraint ordering} $\constrainttorder(\constraints,\dependencies)$ uses the DAG \dependencies in Figure \ref{fig:learning_order} in the following way: all constraints are generated in the topological order of \dependencies. Also, if there is a solid edge from $C_1$ to $C_2$, and $X$ is not a solution to $C_1$, then $X$ is not generated as a candidate group for $C_2$. If there is a dashed edge from $C_1$ to $C_2$ and $X$ is a solution to $C_1$, then $X$ is not generated as a candidate group for $C_2$.

\paragraph{Candidate group generation} $\generategroups(\textit{Constraint,GroupSet,Solutions})$ is the function generating tuples of groups that are legitimate solution candidates e.g., $X,Y$ are variables of \textit{Y = RANK(X)} and $X,Y$ satisfy the constraints from \CSignature above. Essentially, the group generation step is a constraint satisfaction problem associated with the specific constraint. However, many constraints have the same candidate generation procedures, e.g., sum, min, max, avg, count, etc.

\paragraph{Subgroup satisfaction search} $\findassignment(\textit{Constraint,Candidates,Solutions})$ is the function looking for the subsets of vectors in the candidates satisfying the constraint. If multiple subsets satisfy the constraint, a maximal is selected. If $G_1,G_2$, associated with the variables $X,Y$ in the constraint \textit{Y = RANK(X)}, are candidates, then \findassignment selects a single vector $y$ in $G_2$ and a single vector $x$ in $G_1$ such that $x$ is ranked by $y$. For example, in Figure \ref{fig:main_example} the group $G = T_1[{:},3{:}8]$ serves as $X$ and $Y$ and $x$ can be picked as $T_1[{:},7]$ and $y$ would be $T_1[{:},8]$.

\samuel{Expand aggregates to max within group, tweak groups if necessary}
\sergey{[23:32:21] Samuel Kolb: And the motivation is that you can make the group smaller if you want to exclude a column from the aggregate that does not belong in it but accidentally gets included}

\paragraph{Implementation note} Essentially for each constraint both \generategroups and \findassignment are separate constraint satisfaction problems. Hence, it is natural to solve them using declarative solvers. Even though each is a separate constraint problem, they are variations of each other and can be modeled using similar techniques. It also makes an extension of the system easy, we can solve new variations by adding, removing or modifying constraints from a similar constraint satisfaction problem. Furthermore, the declarative languages such as ASP \cite{whaisasp} and Minizinc \cite{minizinc} provide primitives that make modeling transparent and generic.

>>>>

\newcommand{\ecrank}[2]{\ensuremath{#1 = \mathit{RANK}(#2)}}
\newcommand{\ecfkey}[2]{\ensuremath{#1 \rightarrow #2}}
\newcommand{\ecalldiff}[1]{\ensuremath{\mathit{ALLDIFFERENT}(#1)}}

@@

\section{Approach to Tabular Constraint Learning}
As we have seen the statement before we assume the set of groups to be given. It is often, however, not the case and a possible approach is to extract the groups from the CSV file directly. In our system, we have implement the group construction step using the type-consistency check. \sergey{Samuel, we need details here, I think} In general, separating data from a header is a task on its own  \cite{header} and goes beyond the scope of this paper.

@@

Now we describe how the problem of Condensed Tabular Constraint Learning can be solved. The algorithm mimics the structure of the problem in the following way: first we need to order constraints in the order of specificity, then for each constraint we need to generate possible candidate groups, and in the next step we find the subgroups satisfying the constraint. After an iteration we accumulate all solutions and proceed to the next constraint.

@@

\samuel{Running example (RANK?)}
\subsection{Constraint ordering} $\constrainttorder(\constraints,\dependencies)$ uses the DAG \dependencies in Figure \ref{fig:learning_order} in the following way: all constraints are generated in the topological order of \dependencies.
Solid edges point from more general to more specific constraints.
Therefore, if there is a solid edge from $C_1$ to $C_2$, then a set of subgroups satisfying $C_2$ must contain the subgroups satisfying $C_1$.

Dashed edges represent the inverse relationship, pointing from specific to more general constraints.
If there is a dashed edge from $C_1$ to $C_2$, then a set of subgroups satisfying $C_2$ may not contain a subset that also satisfies $C_1$.

\paragraph{Example}
Consider the \ecfkey{FK}{PK} (foreign-key) constraint: for every solution the subgroup (vector) $G_{PK}$ assigned to $PK$ must fulfill \ecalldiff{G_{PK}}.
Therefore, we say indicate \textit{foreign-key} as being dependent on \textit{all-different} and the latter will be treated first.
The \ecrank{Y}{X} constraint does not have any dependencies, however in a restricted version without ties the vector $Y$ would always be a permutation of numbers $1, 2, ..., length(Y)$.

\subsection{Candidate group generation}
$\generategroups(\textit{Constraint,GroupSet,Solutions})$ is the function generating tuples of groups that may contains subgroups satisfying the constraint.
The goal of this step is to prune out impossible groups from the beginning by looking at their properties rather than their actual content (\CSignature vs \CFunction).

Finding candidate groups for an excel constraint~$c$ can be seen as a Constraint Satisfaction Problem (CSP) dependent on the \CSignature of $c$.
Several constraints (e.g. sum, max, count), however, have the same candidate generation procedure as they share the same \CSignature.

\paragraph{Example}
The \ecrank{Y}{X} constraint requires $Y$ to be integer or textual and $X$ to be numeric.
This allows to filter candidate groups for these variables.
Additionally, $Y$ and~$X$ must be vectors of the same length, therefore, only groups with the same length are considered.
Groups that does not satisfy these requirements cannot contain subgroups (vectors) that satisfy \ecrank{Y}{X}.

\subsection{Subgroup satisfaction search}
$\findassignment(\textit{Constraint,Candidates,Solutions})$ is the function that searches for subsets of vectors in the candidate groups that satisfy the constraint.

\paragraph{Example}
Consider a candidate assignment $G_X, G_Y$ associated with the variables $X,Y$ of \ecrank{Y}{X}, then \findassignment selects single vectors $G_X^{i}, G_Y^{j}$ such that \ecrank{G_Y^{j}}{G_X^{i}} is true.
For example, in Figure~\ref{fig:main_example} the group $G = \range{T_1}{\rangeall}{\rangeto{3}{8}}$ can serve as both $G_X$ and $G_Y$.
\findassignment could then select $G_X^{i}$ to be $\range{T_1}{\rangeall}{7}$ and $G_Y^{j}$ to be $\range{T_1}{\rangeall}{8}$ which forms a valid solution.
\\\\
For some constraints (most notably aggregates) there may be multiple overlapping subsets that satisfy the constraint.
We chose to omit subsets that are fully contained within a larger subset since this can lead to overfitting.
If the returned subsets are too large and accidentally contains unrelated vectors, the user can split the group to reflect this structure.

\subsection{Implementation note} Essentially for each constraint both \generategroups and \findassignment are separate constraint satisfaction problems. Hence, it is natural to solve them using declarative solvers. Even though each is a separate constraint problem, they are variations of each other and can be modeled using similar techniques. It also makes an extension of the system easy, we can solve new variations by adding, removing or modifying constraints from a similar constraint satisfaction problem. Furthermore, the declarative languages such as ASP \cite{whaisasp} and Minizinc \cite{minizinc} provide primitives that make modeling transparent and generic.
